Fusion Moves for Correlation Clustering

### 2014-11-06 paper/egpaper_for_review.tex
relatex work in progress

In the present work we present a novel approach that uses the concept of fusion for correlation clustering.
Starting with an initial segmentation we fuse the current segmentations with a proposed one.
Due to the ambiguity of node-labeling for unsupervised segmentations~\cite{} 
original fusion moves~\cite{Lempinski-2010} are improper, see Sec.~\ref{sec:fusion} for discussion.
To overcome this problem we suggest to use fusion moves on the edge-variables that indicates a cut and form a
bijective representation of segmentations.
The subproblem of fusing of segmentations reduce to a much smaller multicut problem which can be solved extremely efficiently.

\textbf{Contribution:}
\begin{inparaenum}[(i)]
\item We propose a fast and scalable move making algorithm for correlation clustering,
\item and give a polyhedral interpretation of this algorithm.
\item We propose two versatile proposal generators, 
\item and evaluate the proposed method on existing and new benchmark problems.
\end{inparaenum}

@@

\section{Notation and Problem Formulation}\label{sec:problem_formulation}
Let $G=(V,E, w)$ be a weighted graph of nodes $V$ and edges $E$.
%
The function $w : E \rightarrow \mathbb{R}$ assigns a weight to each edge.
A positive weight expresses the desire that two adjacent nodes should
be merged, whereas a negative weight indicates
that these nodes should be separated into two different regions.
%
A \emph{subgraph} $G_A = \{A, E_A, w\}$ consists
of nodes $A \subseteq V$ and edges $E_A := E\cap (A\times A)$.
%
A segmentation of the graph $G$ can by either given by an 
node labeling $l \in \mathbb{N}^{|V|}$
or an edge labeling $y \in\{0,1\}^{|E|}$.  
An edge labeling is only consistent if it contains no dangling edges~\cite{Kappes-2013-multicut}.
We denote the set of all consistent edge labelings by $P(G)$.
The convex hull of this set is known as the \emph{multicut polytope} $MC(G)$.

@@

Given a weighted graph $G=(V,E,w)$ we consider the problem of segmenting $G$ such that the costs
of the edges between distinct segments is minimized. This can be formulated in the node~\ref{eq:nodeproblem} 
or edge domain~\ref{eq:edgeproblem}

\begin{eqnarray}
  l^* &=& \argmin_{L \in \mathbb{N}^{|V|}} \sum_{ (i,j) \in E } w_{ij} \cdot [l_{i} \neq l_{j}] \label{eq:nodeproblem}\\
  % y_{ij}^* &=& [l_{u} \neq l_{v}] 
  y^* &=& \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:edgeproblem}%\\ 
\end{eqnarray}

@@

\section{Related Work}\label{sec:related_work}

\subsection{Correlation Clustering}\label{sec:cc}
A major branch of research has focused on solving 
relaxations of eq.~\ref{eq:edgeproblem} by linear programming
or solving  eq.~\ref{eq:edgeproblem} directly to global optimality
by using integer linear programming in a cutting plane fashion  \cite{kappes_2011_emmcvpr,andres_2011_iccv}.

Yarkony et al ~\cite{yarkony_2012_eccv} uses 
a dual decomposition based approach for planar graphs
where the subproblems are planar max cut instances.

@@

\subsection{Fusion Moves}\label{sec:fm}
Move making algorithms, in particular fusion moves, 
have become increasingly popular for energy minimization~\cite{???,kappes_2014_ws}.
For many large scale computer vision applications fusion moves lead to good approximations
with state of the art any time performance~\cite{kappes_2014_ws}.
The fusion move algorithm works as follows:
A current best solution is maintained and proposal solutions 
are generated.
Within each move, the variables can choose the label from the current best or from the 
proposed solution. This problem can be formulated as an 
optimization problem with a binary label space.
Any optimal solution of this auxiliary problem is guaranteed
to not worsen the current best solution.

@@

Given the intrinsic greediness of hierarchical clustering, 
we do not expect very good solutions, but still a good starting 
point for move making algorithm as CGC \cite{beier_2014_cvpr}.
Therefore we propose to initialize  CGC with EBHC
and hope speedup of CGC.

@@

>>>>

\textbf{Contribution:}
In the present work we present some novel approaches that are designed for large scale correlation clustering problems.
First, we define a novel energy based agglomerative clustering algorithm that monotonically increase the energy.
With this at hand we show how to improve the anytime performance of Cut, Clue \& Cut.
Second, we introduce cluster-fusion moves, which extend the original fusion moves~\cite{Lempinski-2010} 
used in supervised segmentation to the unsupervised case and give a polyhedral interpretation of this algorithm.
We propose two versatile proposal generators, and evaluate the proposed methods on existing and new benchmark problems.
That experiments show that we can improve the computation time by one to two magnitudes without worsening the segmentation 
quality significantly.

@@

\section{Notation and Problem Formulation}\label{sec:problem_formulation}
Let $G=(V,E, w)$ be a weighted graph of nodes $V$ and edges $E$.
%
The function $w : E \rightarrow \mathbb{R}$ assigns a weight to each edge.
A positive weight expresses the desire that two adjacent nodes should
be merged, whereas a negative weight indicates
that these nodes should be separated into two different regions.
%
A \emph{subgraph} $G_A = \{A, E_A, w\}$ consists
of nodes $A \subseteq V$ and edges $E_A := E\cap (A\times A)$.
%
A segmentation of the graph $G$ can by either given by an 
node labeling $l \in \mathbb{N}^{|V|}$
or an edge labeling $y \in\{0,1\}^{|E|}$.  
An edge labeling is only consistent if it contains no dangling edges~\cite{Kappes-2013-multicut}.
We denote the set of all consistent edge labelings by $P(G)\subset\{0,1\}^{|E|}$.
The convex hull of this set is known as the \emph{multicut polytope} $MC(G) = \textrm{conv}(P(G))$.

@@

Given a weighted graph $G=(V,E,w)$ we consider the problem of segmenting $G$ such that the costs
of the edges between distinct segments is minimized. This can be formulated in the node domain
by assigning each node $v$ a label $l_v \in \mathbb{N}$
\begin{align}
  l^* &= \argmin_{L \in \mathbb{N}^{|V|}} \sum_{ (i,j) \in E } w_{ij} \cdot [l_{i} \neq l_{j}], \label{eq:nodeproblem}
\end{align}  % y_{ij}^* &=& [l_{u} \neq l_{v}] 
or in the edge domain, by label each edge $e$ as cut $y_e=1$ or uncut $y_e=0$ 
\begin{align}
  y^* &= \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:edgeproblem}.%\\ 
\end{align}
As shown in~\cite{Kappes-2013-multicut} both problems are equivalent, but formulation \ref{eq:nodeproblem}
suffers from ambiguities in the formulation~\cite{Kappes-2011}.

@@

\section{Related Work}\label{sec:related_work}%
%
Due to the ambiguity of formulation~\ref{eq:nodeproblem},
a major branch of research has focused on solving 
relaxations of eq.~\ref{eq:edgeproblem}.
To keep the objective and system of inequalities small, they 
work on sparse graphs and
use cutting plane methods
in combination with (integer) linear programming and efficient 
separation procedures~\cite{kappes_2011_emmcvpr,andres_2011_iccv,kappes_2013_arxiv}.

With no time restrictions and integer constraints these methods can solve the problem to global optimality.
For huge problems both, separation and solving the ILP in each round, becomes very time consuming.

For planar problems, Yarkony \etal~\cite{yarkony_2012_eccv} 
suggested a column generating strategy for the LP-relaxation
that includes all cycle-inequalities of problem \ref{eq:edgeproblem}.
The column generation base on solving planar max cut instances.
While this method is fast, 
it only solve a relaxation of the problem and so requires additional rounding strategies,
lacks of a practical stopping condition for the column generation,
and most important  is restricted to planar problems.

@@

%\subsection{Fusion Moves}\label{sec:fm}
Move making algorithms, in particular fusion moves, 
have become increasingly popular for energy minimization~\cite{???,kappes_2014_ws}.
For many large scale computer vision applications fusion moves lead to good approximations
with state of the art any time performance~\cite{kappes_2014_ws}.
The fusion move algorithm works as follows:
A current best solution is maintained and proposal solutions 
are generated.
Within each move, the variables can choose the label from the current best or from the 
proposed solution. This problem can be formulated as an 
optimization problem with a binary label space.
Any optimal solution of this auxiliary problem is guaranteed
to not worsen the current best solution.

@@

Given the intrinsic greediness of hierarchical clustering, 
we do not expect very good solutions, but still a good starting 
point for move making algorithm as CGC \cite{beier_2014_cvpr}.
Therefore we propose to initialize  CGC with EBHC
and hope to speedup of CGC with this starting point.

@@

Improving proposals.
